# Kagiso Nhlapo

## Data Cleaning

### Task:

1)  Clean biodiversity data

2)  Conduct exploratory analysis and provide some statistics and graphs

3)  Create a country level summary of the data (eg. the total number of species per country)

#### work flow:

-   load packages

-   read in the biodiversity data

-   merge the sheets into one sheet with just two columns

-   Create count of species per country column

-   The data will have : country column, where each row is a country and species count column

-   Create histograms to visualize how species numbers change between countries

-   Do an anova to test if there are difference in countries \*

### 1) Cleaning fo the data

loading libraries

```{r}

library(tidyverse)
library(readxl)
```

Reading the data

```{r}

Bio.dat <- read_xlsx("biodiversity_data.xlsx")

print(Bio.dat)

```

Access all the sheets from the Bio.dat

```{r}

All_sheets <- excel_sheets(Bio.dat)

print(All_sheets)

```

Read all the sheets names into a list of data frames

```{r}

New_data = lapply(setNames(All_sheets, All_sheets),  
                    function(x) read_excel("biodiversity_data.xlsx", All_sheets = x)) 

```

Now combine all sheets into one long data frame

```{r}

combined_data <- bind_rows(New_data)

head(combined_data)


```


### 2)  Conduct exploratory analysis and provide some statistics and graphs

how the data looks per country (Plants and mammals)

```{r}

```

