---
title: "Conservation Alpha Exercise"
author: "Kagiso Nhlapo"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# Conservation Funding Allocation Project

## Part One - Data cleaning & visualization

Loading libraries

```{r}
library(tidyverse)
library(sf)
library(terra)
library(grid)
library(readxl)
library(rio)
library(corrplot)


```

Reading the data biodiversity data from the repository

```{r}

Bio_data <- import_list("/Data/Technical-Assessment-for-Data-and-GIS-Technician//Data/biodiversity_data.xlsx",
                                   setclass = "tbl", rbind = TRUE)

print(Bio_data)

```

Group the biodiversity data by country. Essentially i am counting the number of species found per country then label it "Diversity".

```{r}

Country_level_data <- Bio_data %>%
  group_by(Country) %>%
  summarise(Diversity = n())

print(Country_level_data)

```

Read The conservation funding CSV file from the repository

```{r}

Funding_data <- read.csv("/Data/African_Conservation_Funding.csv")

head(Funding_data)

```

Look at the names of the countries and if they spelled out properly

```{r}
Funding_data$Country %>% unique()
```

Ethiopia is spelled incorrectly, so we need to correct it.

```{r}
Funding_data$Country <- gsub(pattern="Ethiopix", replacement="Ethiopia", Funding_data$Country)
```

Check if its fixed

```{r}
Funding_data$Country %>% unique()
```

Both datasets (funding data and biodiversity data) include a consistent 'country' name column. I then performed a right join of the funding and country-level data based on the country name. This handles the duplicates in the country name column. As a result, we have data for 53 countries (Only DRC is missing for both data sets).

```{r}

Funding.Diversity <- right_join(Country_level_data, Funding_data, by = "Country")

print(Funding.Diversity)

```

Visualization of biodiversity across African countries shows that Ethiopia, Libya, and Cabo Verde have exceptionally high diversity, with values greater than 60 (n \> 60).

```{r}
ggplot(Funding.Diversity, aes(x = Country, y = Diversity)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Visualization of funding for African countries shows that Sudan and Cape Verde have the highest funding, close to USD 20M. Negative funding values are observed for Burkina Faso and Nigeria, while 'N/A' is recorded for Tunisia.

```{r}

ggplot(Funding.Diversity, aes(x = Country, y = Funding..Million.USD.)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The negative values are a source of concern. After visualizing the distribution of the funding data, it became clear that cleaning the data to make all values positive was necessary.

```{r}

hist(Funding.Diversity$Funding..Million.USD.)

```

```{r}

Funding.Diversity$Funding..Million.USD. <- ifelse(Funding.Diversity$Funding..Million.USD. < 0, abs(Funding.Diversity$Funding..Million.USD.),Funding.Diversity$Funding..Million.USD.)

Funding.Diversity$Funding..Million.USD.[is.na(Funding.Diversity$Funding..Million.USD.)] <- 0


head(Funding.Diversity)

```

Lets visualize conservation funding changes have been made.

Histogram:

```{r}
hist(Funding.Diversity$Funding..Million.USD.)
```

Bar graph

```{r}
ggplot(Funding.Diversity, aes(x = Country, y = Funding..Million.USD.)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Part Two - Spatial data cleaning, collation and analysis

Loading the shapefile data for African countries, removing rows with missing ISO3 (three-letter country codes), and retaining the columns needed for analysis and mapping.

```{r}
afriSHP <- st_read("/Data/african_countries_borders/african_countries_borders.shp") %>%
  filter(!is.na(ISO3)) %>%
  dplyr::select(ADM0_CODE,ADM0_NAME,Shape_Leng, Shape_Area,geometry)

print(afriSHP)

```

Set coordinate reference system to WGS 84

```{r}
afriSHP <- st_transform(afriSHP, crs = 4326)

```

I noticed that Cape Verde, Congo, Eswatini, and Ivory Coast are labeled using their former names in the shapefile. It’s important to update these names, as they need to match exactly with the country names used in all other datasets.

```{r}

afriSHP$ADM0_NAME <- gsub(pattern = "Cape Verde", 
                          replacement = "Cabo Verde", 
                          gsub(pattern = "Republic of Congo",
                               replacement = "Congo", 
                               gsub(pattern = "Swaziland",
                                    replacement = "Eswatini",
                                    gsub(pattern = "Côte d'Ivoire",
                                         replacement = "Ivory Coast", 
                                         afriSHP$ADM0_NAME))))


```

Read the Raster for Intactness Index and plot the raster to see how it looks.

```{r}
RAST <- rast("/Data/bii4africa_all.tif")

plot(RAST)
```

Seems like it has a CRS but we will also set it to WGS84

```{r}

crs(RAST)
```

We don't have values for some North African countries, so I expect N/As for the Intactness Index in that region. I processed the data by extracting Intactness Index values using the Africa borders shapefile.

```{r}

II <- terra::extract(RAST, afriSHP)

head(II)

tail(II)

```

Since each country has 30 150 pixels, we need to calculate a single Intactness value for each country. To do this, I aggregated the pixel values for each country.

```{r}

II <- aggregate(bii4africa_all ~ ID, data = II, mean)

print(II)

```

We have 43 columns representing the countries with Intactness values. First, we add an ID column to link the Intactness data with the African countries shapefile. Then, we join the Intactness Index to the shapefile, incorporating the diversity and funding data.

```{r}
afriSHP$ID <- 1:54

afriSHP <- left_join(afriSHP,II, by = "ID")
```

I Perform a left join by country to create a complete dataset with the relevant columns of interest.

```{r}

full_data <- left_join(afriSHP, Funding.Diversity, by = c("ADM0_NAME"="Country"))

print(full_data)
```

Now we keeping all the columns we want to work with and make all N/A = 0

```{r}
full_data <- full_data %>% dplyr::select(ADM0_NAME, Diversity, Funding..Million.USD., bii4africa_all, Shape_Area, geometry)

full_data[is.na(full_data)] <- 0

print(full_data)
```

### Analysis

What is the relationship between funding , intactness and diversity. Are they related linearly ? (Maybe add shape area which is sometimes important in allocating convervation funding)

```{r}
Mod.1_cor <- full_data %>% 
dplyr::select(Diversity, bii4africa_all, Shape_Area, Funding..Million.USD.)%>%
  st_set_geometry(NULL)

M<-cor(Mod.1_cor)

head(round(M,1))

cor <- corrplot(
  M,
  method = "number",
  tl.cex = 0.7,       
  tl.col = "black",
  addCoef.col = "black",  
  number.cex = 0.7,       
  col = colorRampPalette(c("#FFFFFF", "#67a9cf", "#ef8a62"))(100),
  addrect = 4, mar = c(0, 0, 0.0, 0)
)
```

#### 1) Areas with high diversity but low funding.

A weak relationship is observed between all variables. lets precede to understand Areas with high biodiversity but low funding ?

```{r}
full_data$Diversity_To_Finding <- full_data$Diversity/full_data$Funding..Million.USD.

 full_data %>%
  arrange(desc(full_data$Diversity_To_Finding)) %>%
  dplyr :: select(ADM0_NAME, Diversity_To_Finding) %>%
   print()

```

Lets visualize this.

```{r}

ggplot(full_data, aes(x = ADM0_NAME , y = Diversity_To_Finding)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

South Sudan and Eritrea have the highest diversity to funding ratio, but it worth noting that Tunisia has high richness and no funding. Lets consider Areas with high biodiversity intactness but low funding.

#### 2) Areas with high intactness but low funding

```{r}
full_data$Intactness_To_Finding <- full_data$bii4africa_all/full_data$Funding..Million.USD.

 full_data %>%
  arrange(desc(full_data$Intactness_To_Finding)) %>%
  dplyr :: select(ADM0_NAME, Intactness_To_Finding) %>%
   print()
```

Again South Sudan and Eritrea have a high intactness to funding ratio. Lets visualize this.

#### 3) Areas with high intactness and diversity but low funding

```{r}
ggplot(full_data, aes(x = ADM0_NAME , y = Intactness_To_Finding)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Do we observe the same sot of pattern if we look at high richness and intactness but at low funding ?

```{r}
full_data$II_Spp_To_Finding <- (full_data$bii4africa_all*full_data$Diversity)/full_data$Funding..Million.USD.

 full_data %>%
  arrange(desc(full_data$II_Spp_To_Finding)) %>%
  dplyr :: select(ADM0_NAME, II_Spp_To_Finding) %>%
   print()
```

Similarly, South Sudan and Eritrea need more funding for the biodiversity they have. Lets visualize them

```{r}

ggplot(full_data, aes(x = ADM0_NAME , y = full_data$II_Spp_To_Finding)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#### 4) Areas with high intactness , diversity and area shape but low funding

Lets consider area. When allocating funding for conservation, it is important to prioritize areas with high species richness and intactness within a small area, as this may offer better ecological value and conservation potential. But we also have to consider areas with little to no funding. Here we look for small areas with high intactness and richness and low funding.

```{r}

 full_data$Richness_to_Area_ratio <- (full_data$Diversity * full_data$bii4africa_all) / full_data$Shape_Area
 
 full_data$FundingRAR <- (full_data$Richness_to_Area_ratio/full_data$Funding..Million.USD.)
 
 
  full_data %>%
  arrange(desc(full_data$FundingRAR)) %>%
  dplyr :: select(ADM0_NAME, FundingRAR) %>%
   print()

```

Gambia and Eritrea are regions where funding is particularly needed relative to their species richness and intactness, considering their small area size. On the other hand, Burkina Faso and Nigeria exhibit relatively high species richness, but lack intactness and have a smaller area. Let's visualize this distribution to better understand these patterns.

```{r}
ggplot(full_data, aes(x = ADM0_NAME , y = FundingRAR)) +
     geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Part Three- Spatial Heatmaps

### Funding Map

```{r}
ggplot() +
  geom_sf(data = full_data, aes(fill = Funding..Million.USD.)) +
  geom_polygon() +
  labs(title = "",
       fill = "Funding\nin Millions ($)",
       x = "Longitude",
       y = "Latitude")+
  scale_fill_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$Diversity)) +  
scale_colour_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$Diversity))+
    geom_sf_text(data = full_data, aes(label =ADM0_NAME),size=3)+
annotation_scale(location = "br", width_hint = 0.5)


```

Saving the funding map

```{r}

Funding.Map = ggplot() +
  geom_sf(data = full_data, aes(fill = Funding..Million.USD.)) +
  geom_polygon() +
  labs(title = "",
       fill = "Funding\nin Millions ($)",
       x = "Longitude",
       y = "Latitude") +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  geom_sf_text(data = full_data[!full_data$ADM0_NAME %in% c("South Africa", "Mozambique", "Sao Tome and Principe", "Mauritius", "Comoros", "Ghana"), ], 
               aes(label = str_wrap(ADM0_NAME, 1)), 
               size = 3.5, fontface = "bold") + 
  annotate(geom = "text", 
          x = 22, y = -32, size = 3.5, fontface = "bold", label = "South Africa") +
  annotate(geom = "text", 
           x = 38, y = -22, size = 3.5, fontface = "bold", label = "Mozambique") +
  annotate(geom = "text", 
           x = 3, y = 1, size = 3.5, fontface = "bold",label="Sao Tome\n and Principe") + annotate(geom = "text", 
          x = 61, y = -21, size = 3.5, fontface = "bold", label = "Mauritius") +
  annotate(geom = "text", 
           x = 1, y = 6, size = 3.5, fontface = "bold", label = "Ghana") +
  annotate(geom = "text",
           x = 48, y = -11, size = 3.5, fontface = "bold", label = "Comoros") +
  theme(text = element_text(size = 20),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        axis.title = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 24, face = "bold", hjust = 0.5))+
  theme_minimal()+
  annotation_scale(location = "br", width_hint = 0.2,
                   line_width = 1,
  height = unit(0.25, "cm"),
  pad_x = unit(0.25, "cm"),
  pad_y = unit(0.25, "cm"))


ggsave("Output/Fig.1.png", Funding.Map, width = 30, height = 35, units = "cm")



```

### Diversity Map

```{r}
ggplot() +
  geom_sf(data = full_data, aes(fill = Diversity))+
  geom_polygon() +
  labs(title = "",
       colour = "Number\nof Species",
       fill = "Number\nof Species",
       x = "Longitude",
       y = "Latitude") +
  scale_fill_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$Diversity)) +  
scale_colour_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$Diversity))+
  geom_sf_text(data = full_data, aes(label =ADM0_NAME),size=3)+
  annotation_scale(location = "br", width_hint = 0.5)
```

```{r}
Diversity.Map = ggplot() +
  geom_sf(data = full_data, aes(fill = Diversity)) +
  geom_polygon() +
  labs(title = "",
       fill = "Species\nRichness",
       x = "Longitude",
       y = "Latitude")  +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  geom_sf_text(data = full_data[!full_data$ADM0_NAME %in% c("South Africa", "Mozambique", "Sao Tome and Principe", "Mauritius", "Comoros", "Ghana"), ], 
               aes(label = str_wrap(ADM0_NAME, 1)), 
               size = 3.5, fontface = "bold") + 
  annotate(geom = "text", 
           x = 22, y = -32, size = 3.5, fontface = "bold", label = "South Africa")+
  annotate(geom = "text", 
           x = 38, y = -22, size = 3.5, fontface = "bold", label = "Mozambique") +
  annotate(geom = "text", 
           x = 3, y = 1, size = 3.5, fontface = "bold", label = "Sao Tome\n and Principe") +
  annotate(geom = "text", 
           x = 61, y = -21, size = 3.5, fontface = "bold", label = "Mauritius") +
  annotate(geom = "text", 
           x = 1, y = 6, size = 3.5, fontface = "bold", label = "Ghana") +
  annotate(geom = "text", 
           x = 48, y = -11, size = 3.5, fontface = "bold", label = "Comoros")+
  theme(text = element_text(size = 3.5),
    legend.key.height = unit(1, 'cm'),
    legend.key.width = unit(1, 'cm'),
    axis.title = element_text(size = 20, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 17),
    axis.text = element_text(size = 20)) +
   annotation_scale(location = "br", width_hint = 0.2,
                   line_width = 1,
  height = unit(0.25, "cm"),
  pad_x = unit(0.25, "cm"),
  pad_y = unit(0.25, "cm"))+
  theme_minimal()


ggsave("Output/Fig.2.png", Diversity.Map , width = 30, height = 35, units = "cm")


```

### Intactness Map

```{r}
ggplot() +
  geom_sf(data = full_data, aes(fill = bii4africa_all))+
  geom_polygon() +
  labs(title = "",
       colour = "Intactness",
       fill = "Intactness",
       x = "Longitude",
       y = "Latitude") +
  scale_fill_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$bii4africa_all)) +  
scale_colour_gradient2(low = "darkblue", mid = "beige", high = "red", midpoint = median(full_data$bii4africa_all))+
  geom_sf_text(data = full_data, aes(label =ADM0_NAME),size=3)+
  annotation_scale(location = "br", width_hint = 0.5)
```

Save the intactness map

```{r}
II.Map = ggplot() +
  geom_sf(data = full_data, aes(fill = bii4africa_all)) +
  geom_polygon() +
  labs(title = "",
       fill = "Intactness",
       x = "Longitude",
       y = "Latitude")  +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  geom_sf_text(data = full_data[!full_data$ADM0_NAME %in% c("South Africa", "Mozambique", "Sao Tome and Principe", "Mauritius", "Comoros", "Ghana"), ], 
               aes(label = str_wrap(ADM0_NAME, 1)), 
               size = 3.5, fontface = "bold") + 
  annotate(geom = "text", x = 22, y = -32, size = 3.5, fontface = "bold",
           label = "South Africa") +
  annotate(geom = "text", x = 38, y = -22, size = 3.5, fontface = "bold",
           label = "Mozambique") +
  annotate(geom = "text", x = 3, y = 1, size = 3.5, fontface = "bold",
           label = "Sao Tome\n and Principe") +
  annotate(geom = "text", x = 61, y = -21, size = 3.5, fontface = "bold",
           label = "Mauritius") +
  annotate(geom = "text", x = 1, y = 6, size = 3.5, fontface = "bold",
           label = "Ghana") +
  annotate(geom = "text", x = 48, y = -11, size = 3.5, fontface = "bold",
           label = "Comoros")+
  theme(text = element_text(size = 3.5),
    legend.key.height = unit(1, 'cm'),
    legend.key.width = unit(1, 'cm'),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    axis.text = element_text(size = 12))+
  annotation_scale(location = "br", width_hint = 0.2,
                   line_width = 1,
  height = unit(0.25, "cm"),
  pad_x = unit(0.25, "cm"),
  pad_y = unit(0.25, "cm"))+
  theme_minimal()


ggsave("Output/Fig.3.png", II.Map , width = 30, height = 35, units = "cm")
```
