# Kagiso Nhlapo

## Data Cleaning

### Task:

1)  Clean biodiversity data

2)  Conduct exploratory analysis and provide some statistics and graphs

3)  Create a country level summary of the data (eg. the total number of species per country)

#### work flow:

-   load packages

-   read in the biodiversity data

-   merge the sheets into one sheet with just two columns

-   Create count of species per country column

-   The data will have : country column, where each row is a country and species count column

-   Create histograms to visualize how species numbers change between countries

-   Do an anova to test if there are difference in countries \*

### 1) Cleaning fo the data

loading libraries

```{r}

library(tidyverse)
library(readxl)
library(rio)
```

Reading the data

```{r}

Bio_data <- import_list("/Data/biodiversity_data.xlsx",
                                   setclass = "tbl", rbind = TRUE)

print(Bio_data)

```


```{r}

Country_level_data <- Bio_data %>%
  group_by(Country) %>%
  summarise(Diversity = n())

print(Country_level_data)
```

Read all the sheets names into a list of data frames

```{r}

barplot(Country_level_data$Diversity)

```

Now combine all sheets into one long data frame

```{r}

funding <- read.csv("C:/Users/NHLKAG001/Videos/Technical-Assessment-for-Data-and-GIS-Technician/Data/African_Conservation_Funding.csv")


```

### 2) Conduct exploratory analysis and provide some statistics and graphs

how the data looks per country (Plants and mammals)

```{r}

full_data <- left_join(Country_level_data, funding, by = "Country")

```
