---
title: "Creating maps"
author: "Kagiso"
date: "`r Sys.Date()`"
output: html_document
---
---
title: ""
output: html_notebook
---

## These are the maps of conservation of African

load the libraries

```{r}
library(tidyverse)
library(sf)
library(terra)
library(grid)
```

load the data

```{r}
afriSHP <- st_read("african_countries_borders.shp")

print(afriSHP)
```

```{r}
RAST <- rast("C:/Users/NHLKAG001/Videos/Technical-Assessment-for-Data-and-GIS-Technician/Data/bii4africa_all.tif")
print(RAST)
```

extract the biodiversity dat

```{r}
biodivRAST <- terra::extract(biodivRAST, afriSHP)
print(biodivRAST)

```

join this data back to the shapefile dataset

```{r}
afriSHP$biodiv <- biodivRAST$biodivRast

print(afriSHP)


```

join the Intactness Index, country and funding to the shapefile

```{r}

Mean_intactness<- aggregate(bii4africa_all ~ ID, data = biodivRAST, mean)

```

# creating map of Funding

```{r}
afriSHP$ID <- 1:59

afriSHP <- left_join(afriSHP,Mean_intactness, by = "ID")

```

# creating map of biodiversity

```{r}
afriSHP
```

# creating map of intactness index

```{r}
I.index = ggplot()+geom_sf(afriSHP, aes(biodiv))

```

# plot all 3 maps ata high resolution 

```{r}
Maps <- ggrid(funding,
              biodiv, 
              I.index, )
```
